---
title: "Age-Sex Specific Incidence Rates, 2004-2022"
execute: 
  echo: false
---

```{r}
#| label: setup
#| message: false

library(tidyverse)
library(plotly)
library(wesanderson)
```

```{r}
#| label: import_data

# Brought downloaded CSV file with data into project through terminal. Turned
# CSV file into data frame. 

ms_data <- read_csv("DataFilterExport.csv", show_col_types = FALSE) 
```

```{r}
#| label: create_bar_graph

# Graphed Age vs Incidence Rate for each Sex for all years. Tried to create
# faceted bar graph using facet_wrap() but decided it would be more valuable to
# see data for each category next to each other. Used Wes Anderson Palettes R
# Package for the colors in the bar graph. Opted not to add title since it was
# the same as the page and subtitle since key takeaways were listed in paragraph
# below. Used minimal theme for clean look. Adjusted the labels on the x-axis so
# they were easy to read.

ms_bar_all <- 
  ms_data |>
  drop_na() |>
  ggplot(aes(x = Age, y = `Incidence Rate`, fill = Sex)) +
  geom_bar(stat = "identity", position = "dodge", width = 0.7) +
  scale_fill_manual(values = wes_palette("Darjeeling1")[c(2, 3, 1)]) +
  labs(x = "Age Group", y = "Incidence Rate (per 100 000)", 
       fill = "Sex") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))  


# Used plotly package to make graph interactive. Added caption to provide source
# of data, and adjusted sizing and placement of caption. Displayed bar graph.

interactive_bar_all <- 
  ggplotly(ms_bar_all, tooltip = c("Age", "Incidence Rate")) |>
    layout(annotations = list(
      text = "Data from Alberta IHDA",
      x = 1, y = -0.25,
      xref = "paper", yref = "paper",
      showarrow = FALSE,
      font = list(size = 12)), 
      margin = list(b = 100))

interactive_bar_all
```


The trends in across all the years of data available seems to follow many of the
same trends as the data in 2022, such as female incidence rates generally being
higher than male incidence rates. In general, people tend to get diagnosed with
MS as adults, most commonly in their early to middle adulthood, indicating that
that age range is around when MS typically develops in adults.