---
title: "Multiple Sclerosis Age-Sex Specific Incidence Rates"
execute:
  echo: false
---

```{r}
#| label: setup
#| message: false

library(tidyverse)
library(plotly)
library(wesanderson)
```

```{r}
#| label: import_data

# Brought downloaded CSV file with data into project through terminal. Turned
# CSV file into data frame. Isolated only data from 2022, the most recent year
# available.

ms_data <- read_csv("DataFilterExport.csv", show_col_types = FALSE) 
ms_data_clean <- ms_data |>
  filter(Year == "2022") 
```

```{r}
#| label: create_bar_graph

# Graphed Age vs Incidence Rate for each Sex. Tried to create faceted bar graph
# using facet_wrap() but decided it would be more valuable to see data for each
# category next to each other. Added error bars using Standard Error from data
# set. Used Wes Anderson Palettes R Package for the colors in the bar graph.
# Opted not to add subtitle since key takeaways were listed in paragraph below.
# Used minimal theme for clean look. Adjusted the labels on the x-axis so they
# were easy to read.

ms_bar <- 
  ms_data_clean |> 
  drop_na() |>
  ggplot(aes(x = Age, y = `Incidence Rate`, fill = Sex)) +
  geom_bar(stat = "identity", position = "dodge", width = 0.7) +
  geom_errorbar(aes(ymin = `Incidence Rate` - `Standard Error`,
                    ymax = `Incidence Rate` + `Standard Error`),
                position = position_dodge(width = 0.7), 
                width = 0.25,  
                alpha = 0.75) +
  scale_fill_manual(values = wes_palette("AsteroidCity1")) +
  labs(x = "Age Group", y = "Incidence Rate (per 100 000)", 
       title = "Age-Sex Specific Incidence Rates in Alberta, 2022", 
       fill = "Sex") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))  


# Used plotly package to make graph interactive. Added caption to provide source
# of data, and adjusted sizing and placement of caption. Displayed bar graph.

interactive_bar <- 
  ggplotly(ms_bar, tooltip = c("Age", "Incidence Rate")) |>
    layout(annotations = list(
      text = "Data from Alberta IHDA",
      x = 1, y = -0.25,
      xref = "paper", yref = "paper",
      showarrow = FALSE,
      font = list(size = 12)), 
      margin = list(b = 100))

interactive_bar
```



Multiple Sclerosis (MS) is a devastating, demyelinating disease of the central
nervous system that affects people of all ages and genders. MS Incidence Rate
data, which is the number of incident (new) cases divided by the total
population at risk per 100 000 people, was sourced from the government of
Alberta's Interactive Health Data Application. The most recent data, from 2022,
was isolated and graphed. The resulting graphical analysis found that Multiple
Sclerosis tends to affect women more often than men, indicating that gender
might be a risk factor for developing MS.




